---
title: "Simple document"
output: github_document
---

I'm an R Markdown document! 

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

I want to improt 'FAS_litters.csv'.

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```



##'select'
```{r}

select(litters_data, group, litter_number)

pull(litters_data, group)
```

Let's get rid of rows "and": &, "or": |
```{r}
filter(litters_data, gd_of_birth == 20)
filter(litters_data, gd0_weight < 21)
filter(litters_data, group %in% c('Con7','Con8'))
filter(litters_data, gd_of_birth == 20, group == "Con7")

drop_na(litters_data, gd0_weight)
```

## mutate
add or change columns
```{r}
a =mutate(litters_data, 
          weight_change = gd18_weight- gd0_weight,
          group = str_to_lower(group))
```

rearrange
```{r}
arrange(litters_data, gd_of_birth, gd0_weight)
```
pipes
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)
```
```{r}
Sys.setenv(LANGUAGE = "en")
```
## Problem 3
Sol:
  
  Let this random variable $X$ has a normal distribution with $\mu=132$ and $\sigma=9.8$.
  
  a)$P(X>137)=P(Z>\frac{x-\mu}{\sigma})=P(Z>\frac{137-132}{9.8})=P(Z>0.51)=0.305$. 
```{r}
1-pnorm(0.51)
```
So, the probability that a randomly selected American male between ages of 20 and 29 to have a systolic blood pressure above 137.0 is 0.305 to three decimal places.


b) By CLT, $P(\bar{X}<129)=P(\frac{\bar{X}-\mu}{\sigma / \sqrt{n}}<\frac{129-\mu}{\sigma / \sqrt{n}})=P(Z<\frac{129-132}{9.8/ \sqrt{40}})=P(Z<-1.936)=0.026$
```{r}
pnorm(-1.936)
```
So, the probability that the sample mean for blood pressure of 40 males between ages of 20 and 29 will be less than 129 is 0.026 to three decimal places.

c) $P(X<x_{0.9})=0.9$, then $x_{0.9}=144.56$
```{r}
qnorm(0.9,132,9.8)
```
So, the 90 quantile of the sampling distribution of the sample mean X for a sample size of 40 is 144.56 to two decimal places.

## Problem 4

Sol:Because we unknown the population standard deviation, the t-distribution should be used.

a) $\bar{X}-t_{n-1,1-\alpha / 2} \frac{s}{\sqrt{n}} \leq \mu \leq \bar{X}+t_{n-1,1-\alpha / 2} \frac{s}{\sqrt{n}}$
  
$75-t_{30-1,1-0.05 / 2} \frac{8}{\sqrt{30}} \leq \mu \leq 75+t_{30-1,1-0.05 / 2} \frac{8}{\sqrt{30}}$
  
$72.01<\mu<77.99$
```{r}
lower = 75 - qt(0.975, df = 29) * 8/sqrt(30)
upper = 75 + qt(0.975, df = 29) * 8/sqrt(30)
c(lower, upper)
```
So, the 95% confidence interval for the population mean pulse rate of young females suffering from fibromyalgia is (72.01, 77.99).

b) This confidence interval means we have 95% confidence to say that the population mean lies somewhere in the interval (72.01, 77.99).

c) $H_{0}: \mu=70 \text { vs } H_{1}: \mu \neq 70$

$\mathrm{t}=\frac{\bar{X}-\mu_{0}}{s / \sqrt{n}}=\mathrm{t}=\frac{75-70}{8 / \sqrt{30}}=3.42$

$t_{n-1,1-\alpha / 2}=2.05$
```{r}
    # test statistic
    tstat = (75 - 70) / (8/sqrt(30))
    
    # critical value
    cv = qt(0.975, df = 30 - 1)
    
    # decision
    ifelse(abs(tstat) > cv, "reject", "fail to reject")
    
```
$\mathrm{t}>t_{n-1,1-\alpha / 2}$, so we reject $H_{0}$.

$i.e.$ we have enough evidence to reject that the mean is equal to 70.


